Learning Auditory Similarities Between Instrumental Playing Techniques

### 2019-10-08 paper/paper.tex
incorporate comments from meetings

On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
Rather, each excerpt must be played back to multiple human listeners.
Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, thus restricting the size of the musical corpus under study.
Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in artificial neural networks.
Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
The simplest method to compute it in practice is by reverse correlation, i.e. by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but incurring cross-term interference \cite{eggermont1993hearing}.
More recently, the STRF was redefined directly as a spectrographic representation of the spike-triggered average, in order to improve readability \cite{klein2000compneur}.

@@

\ml{introduce the system globally, while stressing the ability of the system to be conditionned for one or many people}

@@

In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects. Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre \ml{vl, citation}, it is  thus taken in this paper as a reasonable assumption. Quantification of this invariance is left for further study.

@@

\ml{provide parameters for scattering
\texttt{
case 'scat'
    filt\_opt.Q = [12 1];
    filt\_opt.J = T\_to\_J(sr*setting.sct/1000, filt\_opt);
case 'tfscat'
    tm\_filt\_opt.Q = [12 1];
    tm\_filt\_opt.J = T\_to\_J(sr*setting.sct/1000, tm\_filt\_opt);
    fr\_filt\_opt.J = 4;}}

@@

In order to better understand the benefits of each design choice, an ablation study \ml{vl ref ??} is now performed using a slightly different protocol. In order to reduce the risk of misleading performance results using learning systems due to overfitting, it common to identify factors of variability that could lead to such problems and mitigate those. In our case, two factors can readily be identified, the variability of the audio data and the one of the perceptual judgments.

@@

The lmnn algorithm is a popular algorithm for solving such a task, but other alternatives can be considered. One of those is the linear discriminant analysis \ml{ja ref ?}\cite{}. When considering the lda in place of the lmnn algorithm, the performance drops down to $81.2\%$.

@@

Considering the separable scattering gives a performance of $93\% \pm 4.2$. \ml{ ja write details on what is lost here ?}

@@

\ml{vl je reecrira le par du dessus en positif fture work}

@@

The recorded instruments belongs to the wind and string classes.
For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
\ja{Better to display these in a table?}

Some instruments can be complemented with sordina (S), wha (W).
The addition of this kind of device modifies the shape of the instrument.
Consequently, an alto saxophone and alto saxophone with sordina are considered as different instruments.
\vl{This is crucial and should not be deferred to the appendix. @writing}
\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}

>>>>

On the other hand, the field of music psychology investigates timbre with the aim of discovering its physiological and behavioral foundations \cite{mcadams2009chapter}.
In this setting, prior knowledge, however accurate, of instrumentation does not suffice to conduct a study.
Rather, each excerpt must be played back to multiple human listeners.
Yet, collecting subjective responses to acoustic stimuli is a tedious and unscalable procedure, thus restricting the size of the musical corpus under study.
Such restriction hampers the applicability of iterative representation learning algorithms, such as stochastic gradient descent in artificial neural networks.
Nevertheless, advanced methods in electrophysiology allow to observe the firing patterns of biological neurons in the presence of controlled stimuli.
This observation, originally carried out on the ferret, has led to a comprehensive mapping of the primary auditory cortex, in terms of its spectrotemporal receptive fields (STRFs) \cite{depireux2001jneur}.
The STRF of a neuron is a function of time and frequency which represents the optimal predictor of its post-stimulus time histogram (PSTH) during exposure to a diverse range of auditory stimuli \cite{aertsen1981biolcyb}.
The simplest method to compute it in practice is by reverse correlation, i.e. by averaging all stimuli that trigger an action potential \cite{deboer1968biomed}.
Historically, STRFs were defined by their Wigner--Ville distribution \cite{flandrin1998book}, thereby sparing the choice of a tradeoff in time--frequency localization, but incurring cross-term interference \cite{eggermont1993hearing}.
More recently, in order to improve readability \cite{klein2000compneur}, the STRF was redefined directly as a spectrographic representation of the spike-triggered average.

@@

\ml{introduce the system globally, while stressing the ability of the system to be conditioned for one or many people (VL). consensus ou un compositeur particulier.}

@@

In order to provide enough data for the numerical analysis described in this section, we assume that the impact of the pitch and nuance on perceptual similarity is much lower than the timbral similarity constraints expressed by the perceptual judgments of the subjects. Assuming invariance to timbre with respect to pitch and nuance is in line with the ansi definition of timbre \ml{vl, citation. reformuler tout le paragraphe}, it is  thus taken in this paper as a reasonable assumption. Quantification of this invariance is left for further study.

@@

\ml{VL. provide parameters for scattering
\texttt{
case 'scat'
    filt\_opt.Q = [12 1];
    filt\_opt.J = T\_to\_J(sr*setting.sct/1000, filt\_opt);
case 'tfscat'
    tm\_filt\_opt.Q = [12 1];
    tm\_filt\_opt.J = T\_to\_J(sr*setting.sct/1000, tm\_filt\_opt);
    fr\_filt\_opt.J = 4;}}

@@

In order to better understand the benefits of each design choice, an ablation study is now performed using a slightly different protocol. In order to reduce the risk of misleading performance results using learning systems due to overfitting, it common to identify factors of variability that could lead to such problems and mitigate those. In our case, two factors can readily be identified, the variability of the audio data and the one of the perceptual judgments.

@@

The lmnn algorithm is a popular algorithm for solving such a task, but other alternatives can be considered. One of those is the linear discriminant analysis \ml{vl ref ?}\cite{}. When considering the lda in place of the lmnn algorithm, the performance drops down to $81.2\%$.

@@

Considering the separable scattering gives a performance of $93\% \pm 4.2$. \ml{ vl. explain separable vs joint. write details on what is lost here ?}

@@

\ml{je reecrirai le par du dessus en positif future work (TODO VL).}

@@

The recorded instruments belongs to the wind and string classes.
For winds, the instruments considered in the second experiment are: accordion (Acc), saxophone alto (ASax), tuba (Tb), bassoon (Bn), clarinet Bb (BbCl), flute (Fl), horn (Hn), oboe (Ob), trombone tenor (TTbn), and trumpet C (TpC).
For the string class, the instruments are: guitar (Gtr), harp (Hp), viola (Va), violin (Vn), violoncello (Vc), and contrabass (Cb).
\ml{Better to display these in a table? VL todo table}
